---
title: "Overview of 4,026 pre-processed gene sets"
author: "Xiang Zhu"
date: "2018-09-16"
output: workflowr::wflow_html
---

[Zhu and Stephens (2018)]: https://www.nature.com/articles/s41467-018-06805-x
[`xiangzhu/rss-gsea`]: https://github.com/xiangzhu/rss-gsea/tree/master/data
[`biological_pathway`]: https://github.com/xiangzhu/rss-gsea/tree/master/data/biological_pathway
[`tissue_set`]: https://github.com/xiangzhu/rss-gsea/tree/master/data/tissue_set
[zenodo-geneset]: https://zenodo.org/badge/latestdoi/55633948

All 4,026 gene sets used in [Zhu and Stephens (2018)][] are freely available at [`xiangzhu/rss-gsea`],
where the folder [`biological_pathway`](https://github.com/xiangzhu/rss-gsea/tree/master/data/biological_pathway)
contains 3,913 biological pathways,
and the folder [`tissue_set`](https://github.com/xiangzhu/rss-gsea/tree/master/data/tissue_set)
contains 113 GTEx tissue-based gene sets.
These gene sets can be referenced in a journal's "Data availability" section
as [![DOI](https://zenodo.org/badge/55633948.svg)][zenodo-geneset].

```zsh
$ tree -L 2 data/
data/
├── README.md
├── biological_pathway
│   ├── gene_37.3.mat
│   └── pathway.mat
└── tissue_set
    ├── de_genes
    ├── he_genes
    └── se_genes

5 directories, 3 files
```

## Biological pathways

The 3,913 public biological pathway used in [Zhu and Stephens (2018)][]
are available in the folder [`biological_pathway`][],
which are represented by two files `gene_37.3.mat` and `pathway.mat`.

The file `gene_37.3.mat` contains basic information of genes.

```
>> load gene_37.3.mat
>> gene
gene =
  struct with fields:
        id: [18732x1 double]
    symbol: {18732x1 cell}
       chr: [18732x1 double]
      desc: {18732x1 cell}
     start: [18732x1 double]
      stop: [18732x1 double]

>> [gene.id(10) gene.chr(10) gene.start(10) gene.stop(10)]
ans =
          18          16     8768444     8878432

>> gene.symbol(10)
ans =
  1x1 cell array
    {'ABAT'}

>> gene.desc(10)
ans =
  1x1 cell array
    {'4-aminobutyrate aminotransferase'}
```

Note that only 18,313 genes mapped to reference sequence were used in our analyses.

```
>> [min(gene.start) min(gene.stop)]
ans =
    -1    -1

>> inref_genes = ~(gene.start == -1 | gene.stop == -1);
>> sum(inref_genes)
ans =
       18313
```

The file `pathway.mat` contains basic information of pathways.

```
>> load pathway.mat
>> pathway
pathway =
  struct with fields:
       label: {4076x1 cell}
    database: {4076x1 cell}
      source: {4076x1 cell}
       genes: [18732x4076 double]
    synonyms: {4076x1 cell}

>> pathway.label(100)
ans =
  1x1 cell array
    {'Activation of NOXA and translocation to mitochondria'}

>> pathway.database(100)
ans =
  1x1 cell array
    {'PC'}

>> pathway.source(100)
ans =
  1x1 cell array
    {'reactome'}
```

The gene-pathway information is represented as a sparse zero-one matrix `pathway.genes`,
where `genes(i,j)==1` if gene `i` is a member of pathway `j` and `genes(i,j)==0` otherwise. 

```
>> genes = pathway.genes;
>> whos genes
  Name           Size                Bytes  Class     Attributes
  genes      18732x4076            3257512  double    sparse

>> genes(:,100)

ans =
      (1243,1)              1
      (3410,1)              1
      (4567,1)              1
      (4668,1)              1  
```

Finally, our analyses only used 3,913 of 4,076 pathways that 

- include 2-499 RefSeq-mapped genes;
- have clear `database` and `source` definitions;
- exclude one pathway `Viral RNP Complexes in the Host Cell Nucleus (PC, reactome)` (because no HapMap3 SNP was mapped to this pathway).

```
>> numgenes = pathway.genes' * inref_genes;
>> size(numgenes)
ans =
        4076           1

>> paths = find(numgenes > 1 & numgenes < 500);
>> size(paths)
ans =
        3916           1

>> database = pathway.database;
>> source = pathway.source;
>> database_na = find(not(cellfun('isempty', strfind(database, 'NA'))));
>> source_na = find(not(cellfun('isempty', strfind(source, 'NA'))));
>> length(union(database_na, source_na))
ans =
     2

>> label = pathway.label;
>> pathway_exclude = 'Viral RNP Complexes in the Host Cell Nucleus';
>> label_include = find(cellfun('isempty', strfind(label, pathway_exclude)));
>> label_exclude = setdiff(1:4076, label_include);
>> label(label_exclude)
ans =
  1x1 cell array
    {'Viral RNP Complexes in the Host Cell Nucleus'}

>> database(label_exclude)
ans =
  1x1 cell array
    {'PC'}

>> source(label_exclude)
ans =
  1x1 cell array
    {'reactome'}
```

## Tissue-based gene sets

The 113 GTEx tissue-based gene sets used in [Zhu and Stephens (2018)][]
are available in the folder [`tissue_set`][].
There are 44 "highly expressed" (HE) gene sets, 49 "selectively expressed" (SE) gene sets
and 20 "distinctively expressed" (DE) gene sets.
The creation of SE sets uses a method described in [Yang et al (2018)](https://doi.org/10.1101/311563).
The creation of DE sets uses a method described in [Dey et al (2017)](https://www.ncbi.nlm.nih.gov/pubmed/28333934).

```{r, engine='bash'}
ls data/tissue_set/he_genes | wc -l
```

```{r, engine='bash'}
ls data/tissue_set/se_genes | wc -l
```

```{r, engine='bash'}
ls data/tissue_set/de_genes | wc -l
```

Each of the tissue-based gene sets has the following format.

```{r, engine='bash'}
head -n 5 data/tissue_set/he_genes/gene_names_expr_LIVER.txt
```

Note that the gene information of tissue-based sets was provided by GTEx,
which may not be the same as `gene_37.3.mat` above.

