---
title: "Overview of 4,026 pre-processed gene sets"
author: "Xiang Zhu"
date: "2018-09-16"
output: workflowr::wflow_html
---

[Zhu and Stephens (2017)]: https://doi.org/10.1101/160770
[`xiangzhu/rss-gsea`]: https://github.com/xiangzhu/rss-gsea/tree/master/data
[`biological_pathway`]: https://github.com/xiangzhu/rss-gsea/tree/master/data/biological_pathway
[`tissue_set`]: https://github.com/xiangzhu/rss-gsea/tree/master/data/tissue_set

```{r,engine='bash'}
tree -L 2 data/
```

## Biological pathways

## Tissue-based gene sets

The 113 GTEx tissue-based gene sets used in [Zhu and Stephens (2017)][]
are available in the folder [`tissue_set`][].
There are 44 "highly expressed" (HE) gene sets, 49 "selectively expressed" (SE) gene sets
and 20 "distincttively expressed" (DE) gene sets.
The creation of SE sets uses a method described in [Yang et al (2018)](https://doi.org/10.1101/311563).
The creation of DE sets uses a method described in [Dey et al (2017)](https://www.ncbi.nlm.nih.gov/pubmed/28333934).

```{r, engine='bash'}
ls data/tissue_set/he_genes | wc -l
```

```{r, engine='bash'}
ls data/tissue_set/se_genes | wc -l
```

```{r, engine='bash'}
ls data/tissue_set/de_genes | wc -l
```

Each of the tissue-based gene sets has the following format.

```{r, engine='bash'}
head -n 5 data/tissue_set/he_genes/gene_names_expr_LIVER.txt
```

